% ----------------------------------------------------------------
% AMS-LaTeX Paper ************************************************
% **** -----------------------------------------------------------
\documentclass[11pt]{amsart}
\usepackage{graphicx, mathabx, amssymb,amsfonts,amsmath,amsthm,newlfont}
\usepackage{epsfig,url}
\usepackage{enumerate,enumitem}
\usepackage[colorlinks=true,linkcolor=red,citecolor=blue]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{color}

\input{aux/commands2}

% ----------------------------------------------------------------

\def\abovespace{\vspace{12pt}}
\def\belowspace{\vspace{8pt}}



\addtolength{\hoffset}{-0.0in} \addtolength{\textwidth}{0in}
\addtolength{\voffset}{-0.0in} \addtolength{\textheight}{0.0in}


% -----------------------------------------------------------------

\title{Term rewriting on nestohedra}

\author{Pierre-Louis Curien}
\address{IRIF, Universit\'e Paris Diderot and $\pi r^2$ team, Inria, France.}
\email{curien@irif.fr}

\author{Guillaume Laplante-Anfossi}
\address{School of Mathematics and Statistics, University of Melbourne, Victoria, Australia.}
\email{guillaume.laplanteanfossi@unimelb.edu.au}

\date{\today}

\subjclass[2020]{Primary 18N20, Secondary 52B11?} 

\keywords{Term rewriting, nestohedra, hypergraph polytopes, categorified operads, categorical coherence, MacLane coherence theorem.}

\thanks{The second author was supported by the Australian Research Council Future Fellowship FT210100256 and the Andrew Sisson Fund.}


\begin{document}

\begin{abstract}
We define term rewriting systems on the vertices and faces of nestohedra, and show that the former are confluent. 
While the associated poset on vertices generalizes Barnard--McConville's flip order for graph-associahedra, the preorder on faces likely generalizes the facial weak order for permutahedra. 
Moreover, we define and study contextual families of nestohedra, whose local confluence diagrams satisfy a certain uniformity condition. 
The proof of confluence of their rewriting systems reproduce, via Huet's correspondence, categorical coherence theorems for monoidal categories, categorified permutads and operads.
\end{abstract}

\maketitle

\setcounter{tocdepth}{1}
%\tableofcontents

\input{sec/introduction2}
\input{sec/nestohedra}
\input{sec/2faces}
\input{sec/rewriting2}
%\input{sec/hyperoperad}
\input{sec/contextual}
%\input{sec/coherence2}
\bigskip

%\emph{Acknowledgements.}   

\bibliographystyle{amsalpha}

\bibliography{Coherence2}


\end{document}



