% ----------------------------------------------------------------
% AMS-LaTeX Paper ************************************************
% **** -----------------------------------------------------------
\documentclass[11pt]{amsart}
\usepackage{graphicx, mathabx, amssymb,amsfonts,amsmath,amsthm,newlfont}
\usepackage{epsfig,url}
\usepackage{enumerate,enumitem}
\usepackage[colorlinks=true,linkcolor=red,citecolor=blue]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{color}

\input{aux/commands}

% ----------------------------------------------------------------

\def\abovespace{\vspace{12pt}}
\def\belowspace{\vspace{8pt}}



\addtolength{\hoffset}{-0.0in} \addtolength{\textwidth}{0in}
\addtolength{\voffset}{-0.0in} \addtolength{\textheight}{0.0in}


% -----------------------------------------------------------------

\title{Topological proofs of categorical coherence}

\author{Pierre-Louis Curien}
\address{IRIF, Universit\'e Paris Diderot and $\pi r^2$ team, Inria, France.}
\email{curien@irif.fr}

\author{Guillaume Laplante-Anfossi}
\address{School of Mathematics and Statistics, The University of Melbourne, Victoria, Australia.}
\email{guillaume.laplanteanfossi@unimelb.edu.au}

\date{\today}

\subjclass[2010]{Primary 18N20, Secondary 52B11} 

\keywords{Categorified operads, categorical coherence, Seifert--Van Kampen theorem, polytopes, MacLane coherence theorem, rewriting theory, Morse theory.}

\dedicatory{"We shall construct $KP_n$, as a CW-complex, in Section 2 and show that it is an $(n-1)$-ball. This gives an instant one-step proof of MacLane's theorem in full generality."  \\ --  Mikhail M. Kapranov}

\thanks{The second author was supported by the Andrew Sisson Fund and the Australian Research Council Future Fellowship FT210100256.}

%\thanks{The second author was supported by the European Union's Horizon 2020 research and innovation program under the Marie Sklodowska-Curie grant agreement No 754362 \inlinegraphics{EU.png}, by the Natural Sciences and Engineering Research Council of Canada (NSERC) and by the ANR-20-CE40-0016 Higher Algebra, Geometry and Topology.}


\begin{document}

\begin{abstract}
\correction{We give a short topological proof of coherence for categorified non-symmetric operads by using the fact that the diagrams involved form the 1-skeleton of simply connected CW complexes. 
We also obtain a ``one-step'' topological proof of Mac Lane's coherence theorem for symmetric monoidal categories, as suggested by Kapranov in 1993.  
Our analysis is based on a notion of combinatorial homotopy, which we further study in the special case of polyhedral complexes, leading to a second geometrical proof of coherence which is very close to Mac Lane's original argument.  
We use Morse theory to show that this second method is (strictly) less general than the first.
%we use Morse theory to give a second topological proof which is very close to Mac Lane's original argument. 
We provide a detailed analysis of how  both methods allow us to deduce these two categorical coherence results and discuss possible generalizations to higher categories. }
\end{abstract}

\maketitle



\setcounter{tocdepth}{1}
%\tableofcontents

\input{sec/introduction}
\input{sec/polycoherence}
\input{sec/catoperads}
\input{sec/monoidal}
\input{sec/applications}
\bigskip

\emph{Acknowledgements.}   
We would like to thank all the participants of the Roberta Seminar held on September 30th, 2020 in Paris for planting the seeds of the present note.  
We would like to thank Andrea Bianchi for enlightening discussions, for pointing out mistakes in a first draft, and for reminding us of the Seifert--Van Kampen theorem. 
Finally, we are grateful to Zoran Petri{\'c} for precious discussions on weak Cat-operads, to Marcy Robertson \correction{and to the anonymous referee} for helpful comments \correction{which led to substantial improvements of the paper}.
\correction{
The first author wishes to thank the Mathematisches Forschungsinstitut Oberwolfach for several visits as a Research Fellow in 2020 and 2021, thanks to which much of the work presented here, as well as in the companion paper \cite{CLA24}, could take off.
The second author would like to thank the Copenhagen Center for Geometry and Topology as well as the Max Planck Institute for Mathematics in Bonn where part of this work was carried out.}


\bibliographystyle{amsalpha}

\bibliography{Coherence}


\end{document}



