% !TEX root = ../Coherence.tex

\section{Perspectives}


\Guillaume{Flow categories?}


\subsection{Rewriting theory and Koszul duality}
Any generic vector $\vec v$ defines a canonical rewriting system on the vertices of a polyhedral complex $X$, which is moreover always terminating and confluent. 
The confluence of the rewriting system described in the second proof of \cref{thm:coherence-operahedra} can be used to give a proof that the colored operad $\mathcal{O}$ \cite[Definition 4.2]{laplante-anfossiDiagonalOperahedra2022a} encoding non-symmetric operads is Koszul (and which avoids resorting to the diamond lemma!): one can proceed with the rewriting method \cite[Section 8.3]{LodayVallette12} adapted to colored operads \cite{KhariKhoro20}, along the lines of the Remark following \cite[Theorem 9.1.5]{LodayVallette12}. 

In the other way around, one can use the fact that $\mathcal{O}$ is Koszul to prove the coherence theorem for categorified non-symmetric operads, using \cite[Theorem 15]{marklCoherenceConstraintsOperads2001} in a suitable generalization of \cite[Example 16]{marklCoherenceConstraintsOperads2001}.
Thus, the categorical coherence theorem for $\mathcal{O}$-algebras and the fact that $\mathcal{O}$ is Koszul are equivalent.
Considering the large scope of the theory developed in \cite{marklCoherenceConstraintsOperads2001}, this could be a general fact. 

\Guillaume{Application to Drinfeld}
\Guillaume{Morse theory and rewriting; polygraphs, higher dimensional rewriting?}

\subsection{Higher categories} 
\label{sec:higher}
\cref{thm:top-coherence} demonstrates that, in the case of monoidal categories, coherence is equivalent to the vanishing of the first homotopy groups of the associahedra. 
Since the associahedra are contractible, and therefore all their homotopy groups vanish, one could hope for a topological proof of higher dimensional coherence theorems.
Seeing a monoidal category as a bicategory with one object, one can ask about a coherence theorem for tricategories with one object. 
A first look at the diagrams in the beginning of \cite[Section 2]{gordonCoherenceTricategories1995} suggests that such a theorem should be at least related to the vanishing of the second homotopy groups of the associahedra.  
To formulate higher dimensional statements, one needs a good structure of pasting scheme on each associahedron, which is the subject of ongoing work \cite{AMMLA}. 

Recent results of S. Barkan provide evidence for these higher dimensional statements, in the context of $\infty$-operads \cite{barkanArityApproximationInfty2022}.
In this vein, it seems likely that the present results could be interpreted as a strict version and a special case of \cite[Theorem B]{barkanArityApproximationInfty2022}. 
It would be interesting to see how the permutoassociahedra arise in the strictification process, and how they are related to operadic partition complexes.  



